### 2022-03-17 InfHyp.Rmd
first commit

---
title: "Bayesian Evidence Synthesis Methods"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Matrix)
library(lavaan)
library(BFpack)
library(ggplot2)
library(metafor)
library(bain)
```

## Context
Caspar developed a method to compute an evidence synthesis BF. To do so, he (more or less) creates a datasets with $k$ independent groups, where each group represents a separate study. With this multigroup dataset, we can use `bain` or `BFpack` to calculate 'overall Bayes factors'. For $BF_{i,u}$ this is the same as the product BF that we use in Bayesian evidence synthesis. For $BF_{i,c}$ it is not. Below, I explain the difference and the link between both methods. The remaining challenge is to decide what is conceptually correct.

## Example Data + Results
In this example our evidence synthesis hypothesis with $k=3$ datasets is $H_i: r_{data1}>0.1 \& r_{data2}>0.1 \& r_{data3}>0.1$.

```{r sim}
#hyperparameters simulation
#es = effect size
#tau2 = between-study variance in es
#participants
#k=number of datasets
hyperparams <- list(
    es = 0.3,
    tau2 = 0,
    pp = 120,
    k = 3
)

#hyp_value = threshold for > in informative hypothesis
hyp_value <- .1

sim_cond <- do.call(expand.grid, hyperparams)

#hyperparameters population
#true_es = effect size cor in population
sim_data <- function(true_es, tau2, pp){
    rho <- rnorm(1, true_es, sqrt(tau2)) #reliability
    x = rnorm(pp, 0, 1) # vector of predictor?
    y = rnorm(pp, rho*x, sqrt(1-rho^2)) #outcome
    list(x = x, y = y)
}

colSD <- function(x) apply(x, 2, sd, na.rm = TRUE)

# generate data
set.seed(6164900)
dfs <- lapply(1:hyperparams$k, function(n){
    sim_data(true_es = hyperparams$es, tau2 = hyperparams$tau2, pp = hyperparams$pp)
})
# Compute correlations and se (correlations between x and y)
res <- sapply(dfs, function(x){
    r = do.call(cor.test, x)
    unname(c(r$estimate, sqrt((1 - r$estimate^2)/r$parameter)))
})
# Prepare for BFpack calls
est <- res[1, ]
names(est) <- paste0("r", 1:length(est))
Sig <- diag(res[2, ])
# Get individual BFs for each sample
bf_individual <- lapply(paste0(names(est), ">", hyp_value),
                        bain,
                        x = est,
                        Sigma = Sig,
                        n = hyperparams$pp)

fit.i <- sapply(bf_individual, function(x){x$fit$Fit[1]}) #fits of hypotheses (posterior probability of hypotheses given the data)
#in this case, posterior probability that rho > .1 given the data in each df.
com.i <- sapply(bf_individual, function(x){x$fit$Com[1]}) #fit of complement given the data.
bf_prod_c <- sapply(bf_individual, function(x){x$fit$BF.c[1]}) #BFic this is the complement and thus it is just fit.i/(1-fit.i)
bf_prod_u <- sapply(bf_individual, function(x){x$fit$BF.u[1]}) #BFiu this is just fit.i/com.i, so posterior prop of h1/posterior prop of hu

# Get overall BF
bf_together <- bain(x = est,
                    hypothesis = paste0("(", paste0(names(est), collapse = ", "), ") > ", hyp_value),
                    Sigma = Sig,
                    n = hyperparams$k*hyperparams$pp)

fit.t <- bf_together$fit$Fit[1]
com.t <- bf_together$fit$Com[1]

results = list(fit.i=fit.i,
               prod.fit=prod(fit.i), #product of informative hypotheses fits
               com.i=com.i,
               prod.com=prod(com.i),
               bf.ic=bf_prod_c,
               bf.ic.prod=prod(bf_prod_c),
               fit.t=fit.t,
               com.t=com.t,
               bf.ct=bf_together$fit$BF.c[1],
               bf.ict=prod(fit.i)/prod(com.i)/((1-prod(fit.i))/(1-prod(com.i))),
               bf.iu.prod=prod(bf_prod_u),
               bf.ut=bf_together$fit$BF.u[1])

print(results)

```

## Explanation
The product of the individual $BF_{i,u}$ = `r results$bf.iu.prod`, this equals the combined data product `r results$bf.ut`.

The product of the individual $BF_{i,c}$ = `r results$bf.ic.prod`, whereas the combined data product is `r results$bf.ct`.

The product Bayes factor equals $\frac{\prod f_i / \prod c_i}{(\prod 1-f_i)/(\prod 1-c_i)}$ = $\frac{`r results$prod.fit` / `r results$prod.com`}{`r prod(1-results$fit.i)` / `r prod(1-results$com.i)`}$ = `r results$bf.ic.prod`.

The combined data product can be recomputed based on the elements of the individual Bayes factors by $\frac{\prod f_i / \prod c_i}{(1-\prod f_i)/(1-\prod c_i)}$ = $\frac{`r results$prod.fit` / `r results$prod.com`}{`r 1-results$prod.fit` / `r 1-results$prod.com`}$ = `r results$bf.ict`.

The difference is in the denominator: should we look at the within-study complement or is it (also) correct to take a complement of the synthesized hypothesis?

With the individual product method, our complement is: "r_data_1<0.1 \& r_data_2<0.1 \& r_data_3<0.1", which has the same complexity as our hypothesis "r_data_1>0.1 \& r_data_2>0.1 \& r_data_3>0.1" = $0.5^k$ = `r 0.5^3`. In the overall method, the complement is 'not Hi', which encompasses more options = $1-0.5^k$ = `r 1-0.5^3`.

The question is whether we first test against the complement in each study and take $BF_{i,c} \times BF_{i,c} \times BF_{i,c}$ or if we can compute the synthesized $BF_{i,c}$ for $H_i: r_{data1}>0.1 \& r_{data2}>0.1 \& r_{data3}>0.1$ vs "not $H_i$".

&nbsp;