### 2022-07-20 manuscript.Rmd
Caspar's edits of the tutorial

---
title             : "Tutorial: Aggregate evidence from heterogeneous replication studies using the product Bayes factor"
shorttitle        : "PRODUCT BAYES FACTOR"

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Formal Analysis
      - Funding acquisition
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Writing – original draft
      - Writing – review & editing
  - name: "Eli-Boaz Clapper"
    affiliation: "1"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Utrecht University, dept. Methodology & Statistics"
  - id            : "2"
    institution   : "Open Science Community Utrecht"

authornote: |
  This is a preprint paper, generated from Git Commit # `r substr(gert::git_commit_id(),1,7)`. This work was funded by a NWO Veni Grant (NWO Grant Number VI.Veni.191G.090), awarded to the lead author.

abstract: |
  TODO

keywords          : "bayes factor, evidence synthesis, bayesian, meta-analysis"
wordcount         : "5356"

bibliography      : ["product_bayes.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
knit: worcs::cite_essential
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
run_everything <- FALSE
library("papaja")
library(tidySEM)
library(kableExtra)
r_refs("r-references.bib")
out <- readRDS("out.RData")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning = FALSE)
```

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
Replication research has come into focus as a solution to this crisis and as a way to derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
<!--For many, highly regarded published studies however, results could not be replicated [@brembs2018prestigious].-->
<!--A lack of validating replication research partly strengthened the belief that single studies results should solely be viewed as part of many studies studying the same research question [@asendorpf2016recommendations]. -->
<!--Evidence over multiple studies must be aggregated to derive more reliable conclusions and within this paradigm, research synthesis methods have blossomed [@shadish2015meta; @nakagawa2019research].-->
In step with this interest in replication research,
research synthesis methods have blossomed [REF].
These methods aggregate the findings of multiple studies,
and thus enable drawing overarching conclusions across multiple studies.
The present paper focuses on quantitative methods for research synthesis [e.g., @vanlissaSelectRelevantModerators2021], although qualitative [e.g., @vanlissaMappingPhenomenaRelevant2021] research synthesis methods also exist.

A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
Heterogeneity appears when, for example, studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations.
The most common quantitative research synthesis method is meta-analysis,
in which results of different studies are aggregated to estimate an aggregate effect size [@borensteinIntroductionMetaAnalysis2009].
In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
First, if studies are exact replications, one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
Second, when heterogeneity between studies can be assumed to be because of random measurement error, random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
And finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators. 
Each of these approaches requires making different assumptions about the nature of heterogeneity.

An alternative approach for evidence synthesis that does not impose such assumptions is Bayesian evidence synthesis (BES) [@kuiperCombiningStatisticalEvidence2013]. 
Instead of aggregating effect sizes, BES aggregates the evidence in favor of an informative hypothesis $H_i$ across studies.
The amount of evidence for this hypothesis is expressed as a Bayes factor, BF.
A Bayes factor can be interpreted as the ratio of evidence in favor of $H_i$ divided by evidence against it.
Thus, BF = 10 means that the data provide ten times more support in favor of the hypothesis than against it.
These Bayes factors can be synthesized across studies by taking their product.
The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
Note that other approaches to BES exist [see "Bayesian Evidence Synthesis" in @heckReviewApplicationsBayes2022].
The only assumption of BES is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.

Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
Meta-analysis estimates the value or distribution of a population effect size. 
It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
It thus answers the question: Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
BES, on the other hand, aggregates evidence in favor of an informative hypothesis across multiple studies.
It thus answers the question: Do all these studies support the informative hypothesis?
Both methods thus provide complementary information.

This tutorial paper introduces the first implementation of BES in user-friendly open source software.
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `r as.character(packageVersion("bain"))`.
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

# Simulation study

The present simulation study set out to validate the PBF algorithm.
For each iteration of the simulation, we simulated a correlation coefficient in multiple samples.
The informative hypothesis was kept constant at $H_i: \rho > .1$.
All simulation conditions were evaluated once in the presence of a true population effect,
defined as $\rho = .2$,
and once in the presence of a null effect, defined as $\rho = .1$.
A PBF > 3 was used as a decision criterion to conclude that $H_i$ was supported.
As a benchmark for comparison, we used several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was a "vote count" algorithm,
which is common but not considered to be good practice [REF].
This approach used one-sided z-tests to examine whether a null hypothesis corresponding to the informative hypothesis was rejected in the majority of samples, i.e.: $H_0: \rho = .1$.
The second was a random-effects meta-analysis (RMA), which is the standard in the field.
For this algorithm, the null-hypothesis was rejected if a 90% confidence interval for the overall effect size excluded $H_0$.
Third was an individual participant data (IPD) meta-analysis.
Like classic meta-analysis, IPD is a multilevel model, clustered by sample.
In contrast to classic meta-analysis, however, IPD freely estimates variance at the first level, because raw data are available.
The PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
It is thus informative to compare the PBF to both these methods.
IPD was also evaluated using a 90% confidence interval for the overall effect size.

## Performance indicators

For each algorithm, a confusion matrix was obtained by tabulating inferential decisions made using the criteria described above against the population status of the hypothesis (true or false).
This matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP) and false negatives (FN).
These quantities were summarized as sensitivity, $1-\frac{FN}{TP + FN}$, the probability of concluding the hypothesis is false given that it was truly false in the population, <!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there. I found that its formula is usually presented as TP/(TP+FN)-->
<!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
and specificity, $1-\frac{FP}{FP + TN}$, the probability of concluding the hypothesis is true given that it was also true in the population. <!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive/true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.

## Design factors

To examine performance in a range of realistic scenarios,
several design factors were manipulated:
The sample size per group $n \in (20, 80, 200, 500)$,
the number of groups $k \in (2, 3, 10)$, where 2 is the minimum number of groups and 10 is a very large number of groups,
and the reliability of the two correlated variables, $\alpha \in (0.6, 0.8, 1.0)$, where 0.6 is the lowest reliability conventionally considered to be acceptable, and 1 represents perfect reliability, as is assumed when analyzing correlations between observed items or scale scores.
The design factors combined to produce `r out$conditions` unique conditions.
For all simulation conditions, 1000 data sets were generated.

```{r, echo = F}
library(data.table) # load in libraries
library(DT)
library(ggplot2)
# rename <- c("prodbf" = "PBF", "gpbf" = "gPBF", "tbf" = "TBF", "ipd" = "IPD", "rma" = "RMA", "allsig" = "VC")

```

# Results

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
# names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
First, we examined overall model performance across conditions.
These results indicate that all algorithms except PBF had low sensitivity to detect a true effect.
In contrast, specificity was very high for all algorithms except PBF.
This suggests that the other algorithms classified most conditions as negatives (no effect found),
regardless of the existence of a population effect.
The PBF trades a loss of specificity for increased sensitivity,
and average levels of both were approximately equal.
As the PBF vastly outperformed other BFs, these were omitted from further analysis.

## Effect of simulation conditions

We examined the effect of simulation conditions on overall accuracy.
```{r tabeffect, echo = FALSE}
tab <- read.csv("effect_of_conditions.csv", stringsAsFactors = FALSE)
tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
tab <- tab[, grepl("^\\w+$", names(tab)) | (grepl("\\.{2}", names(tab)) & grepl("PBF", names(tab)))]
names(tab)[grepl("\\.{2}", names(tab))] <- paste0("vs ", gsub("(\\.|PBF|vs)", "", names(tab)[grepl("\\.{2}", names(tab))]))
knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
PBF performance was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
<!-- should we also test for interaction effects? I seem to remember that there were some interesting effects of errorsd*k and n*k -->
<!-- CJ: I don't think so, that might be more relevant for one of the papers that focused on the simulation study. For this tutorial, a simpler sim is fine. -->

### Effect of sample size

```{r fign, fig.cap="Mean performance by sample size"}
res <- readRDS("confusion_by_cond.RData")

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("n", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Sample size")
```

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increase with sample size.
The other algorithms also show increasing sensitivity, but not specificity, which is at a ceiling.
This difference explains the effect of reliability on the diference between algorithms (see Table \@ref(tab:tabeffect)).

### Effect of number of groups

```{r, fig.cap="Mean performance by number of groups"}
df_plot <- data.table::melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("k", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Number of groups")
```
The figure indicates that, for PBF at higher levels of k, lower sensitivity is exchanged for greater specificity.
The other algorithms do not show this pattern, as their specificity is at a ceiling.
This difference in pattern of effects explains why number of groups has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Only VC shows decreasing sensitivity with an increasing number of groups;
this is because the probability of obtaining any false negatives increases with the number of groups.

### Effect of reliability

```{r, fig.cap="Mean performance by reliability"}
df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("reliability", "Outcome", "alg")]
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("reliability", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
ggplot(df_plot, aes(x = reliability, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  labs(y = "Mean performance", x = "Reliability")
```

The figure indicates that for PBF, at higher reliability, lower sensitivity is exchanged for greater specificity.
The other algorithms do not show this pattern, as their specificity is at a ceiling.
Their sensitivity decreases with lower reliability,
and therefore, so does their overall performance.
This difference in pattern of effects explains why reliability has a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Note that, in contrast to other algorithms, the performance of PBF is less susceptible to reliability.

# Discussion

Within the scope of this simulation, PBF had relatively better inferential properties than the other algorithms under consideration.
Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.
<!-- Conventional threshold for type 1 error: 5% -->
<!-- Conventional threshold for power: 80% -->
<!-- J. Cohen, Statistical Power Analysis for the -->
<!--  Behavioral Sciences, 2nd ed. (Erlbaum, Hillsdale, -->
<!--  NJ, 1988). This is the source of the system of power -->
<!--  analysis described here; the power values and sam -->
<!--  ple sizes of the illustrations derive from this book's -->
<!--  tables. -->
<!--  2. J -->

An important caveat is that none of the algorithms met conventional criteria for power (80%) and type I error (5%).
Overall, PBF had the best performance, with power of 76% and type I error of 24%.
Thus, if researchers intend to synthesize evidence for an informative hypothesis across heterogeneous studies, PBF may be the most suitable method - but its limitations should be acknowledged in applied research.

The overall performance of PBF increased most with increasing sample size.
With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.
With increasing reliability, increasing specificity was traded off for decreasing sensitivity.

# Tutorial

This tutorial about Bayesian Evidence Synthesis (BES) requires R-package `bain` version `r packageVersion('bain')` or later.
This tutorial focusses on synthesizing evidence for an informative hypotheses using the Product Bayes Factor (pbf). 
The data used to illustrate the method have been simulated based on the data presented in [@vanleeuwen2022morality].
These simulated data are available in the `bain` package. 
A more detailed description of the dataset is either provided in [@vanleeuwen2022morality] or by running `?bain::synthetic_us`, `?bain::synthetic_dk` or `?bain::synthetic_nl`.
The research question is whether there is an association between political orientation and moral dispositions.
Suitable data were collected in three countries: the United states of America, Denmark and the Netherlands.
Each sample contains multiple measures of political orientation and moral dispositions.
In the original publication, the PBF was used to aggregate evidence across scales and countries to obtain an overall measure of support for the central hypothesis.
This tutorial follows the same rationale, but uses only one effect size per sample,
and varies the way this effect size is computed to illustrate different use cases.

```{r}
NL <- bain::synthetic_nl
DK <- bain::synthetic_dk
US <- bain::synthetic_us
```

<!-- skeleton
1.) explain calculation of single BF and what it means, refer to previous tutorial on the bayes factor
2.) introduce the case where there are multiple studies and we want to evaluate a single hypothesis
3.) introduce pbf(), its uses, outputs and interpretation
-->

### How to use bain

<!-- CJ: Is dit niet teveel herhaling van wat we in de inleiding zeggen?? -->
<!-- In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis.  -->
<!-- This method evaluates a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data.  -->
<!-- However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis. They may also not only be interested in whether the hypothesis is probable, but also by how much.  -->
<!-- In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis.  -->
<!-- Because the focus of this tutorial is on BES,  -->
First, we briefly introduce the basic use of `bain::bain()`, and how to interpret its output. 
Suppose we are interested in whether the self-reported importance of family morality positively predicts a conservative socio-political orientation.
Because both scales consist of multiple items, we use structural equation modeling (SEM) to perform latent variable regression.
Our hypothesis is $H_i: \beta > 0$, where $\beta$ (beta) is the standardized regression coefficient.
Note that, for the sake of simplicity, this hypothesis corresponds to a traditional one-sided null-hypothesis test.
The original publication instead used a minimal effect size of interest, and specified $H_i: \beta > .1$.
The code below illustrates the analysis for a single sample.
For a more in-depth tutorial on `bain::bain()`, see @hoijtink2019tutorial, and for further guidance on the use of `bain::bain()` in SEM, see @REF VAN LISSA TUTORIAL.

<!--EB: For simplicities sake, I for now create simple average scores without taking into account reverse coding. This should likely be improved with factor scores -->
<!-- CJ: Keep the code SUPER simple. No functions etc -->

```{r echo = TRUE, eval = FALSE}
# Load relevant packages
library(bain)
library(lavaan)

# Specify a SEM-model with two latent variables and latent regression
model_nl <- "
family =~ fam_1 + fam_2 + fam_3
conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
conservative ~ beta * family"

# Estimate the model in lavaan
results_nl <- sem(model = model_nl, data = NL)

# Test the hypothesis that the effect size labeled 'beta' is positive
# The argument standardize = TRUE uses the standardized model parameters
bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
```
```{r echo = FALSE, eval = TRUE}
# Load relevant packages
library(bain)
library(lavaan)

# Specify a SEM-model with two latent variables and latent regression
model_nl <- "
family =~ fam_1 + fam_2 + fam_3
conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
conservative ~ beta * family"

# Estimate the model in lavaan
results_nl <- sem(model = model_nl, data = NL)

# Test the hypothesis that the effect size labeled 'beta' is positive
# The argument standardize = TRUE uses the standardized model parameters
bf_nl <- bain(results_nl, hypothesis = "beta > 0", standardize = TRUE)
bf_nl
```

<!-- Socio-political orientation and group loyalty have been aggregated to variables `soc` and `grp` respectively.  -->
<!-- Both are continuous variables with higher scores indicating either conservatism or higher group loyalty.  -->
<!-- In NHST we specify the null hypothesis that there is no association between `soc` and `grp`.  -->
<!-- One method to test the hypothesis is to calculate a standardized regression coefficient. -->

<!-- $H_0: \beta_{grp} = 0$. -->

<!-- However, we may have prior expectations that higher group loyalty is somewhat related to  higher conservatism.  -->
<!-- An informative hypothesis could be: -->

<!-- $H_i: \beta_{grp} > 0$ -->

<!-- SESAMESIM DATA: Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`). `postnumb` is a continuous variable which ranges from 0-63. `site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking. In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  site`:

$H_0: \mu_{site1} =\mu_{site2} = \mu_{site3} = \mu_{site4} =\mu_{site5}$

However, we may have prior expectations that the children growing up in an advantaged site will have a higher score than children growing up in a disadvantaged site. Our informative hypothesis is:

$H_i: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$ -->

<!-- Because it is very improbable that $\beta_{grp}$ is exactly equal to 0, more information can be won by testing $H_i$. -->
<!-- We can test $H_i$ by running a linear regression on the data and passing the `lm` object with the specific hypotheses to `bain`.  -->
<!-- For a single study, such as the study that investigated the Netherlands, the Bayes factor would be obtained as follows: -->

```{r, echo = FALSE, eval = FALSE}
# load bain package
library(bain)

# specify model, allow meaningful zero-point
fit <- lm(soc ~ grp, data = NLa)

#specify Hypotheses to evaluate
hypotheses <- "grp > 0"

fit_bain <- bain(fit, hypothesis = hypotheses)
print(fit_bain)

```
<!-- SESAMESIM DATA: The output of `bain` seems to support neither hypothesis. The `fit` for both hypotheses is approximately zero which means that almost none of the data are in agreement with the restrictions the specified hypotheses pose. The `BF.c`, which is the Bayes Factor of the hypothesis versus its complement, is for both hypotheses close to zero. This implies that the complement of both hypotheses is $\frac{1}{BF.c}$ times more likely than the passed hypotheses. Finally, `PMPc`, the Posterior Model Probabilities of the specified hypotheses with $H_c$ included, is also close to zero for both specified hypothesis. This means that *out of* the specified hypotheses and their complement, the probability of $H_c$ being true is 99.9%. -->

Note that the hypothesis is strongly supported when compared to its complement, $BF_c `report(bf_nl$fit$BF.c[1])`$.

### Aggregating evidence across studies

As mentioned before, suitable data were collected to test the substantive hypothesis in three countries.
There are differences between countries that prevent analyzing these data as a multilevel model, however.
For instance, conservatism was measured using different scales.
This is an appropriate situation to use the PBF to aggregate evidence across countries.
First, we estimate a latent regression model for the remaining two countries, taking care to label the parameter of interest with the same label in all samples.
Then, we bind all three SEM-models in a list, and call PBF to evaluate the hypothesis of interest on all models and aggregate the evidence.

```{r, eval = FALSE, echo = TRUE}
# Specify the models for DK and US
model_dk <- "
family =~ fam_1 + fam_2 + fam_3
conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
conservative ~ beta * family"
model_us <- "
family =~ fam_1 + fam_2 + fam_3
conservative =~ secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
                secs_soc_6 + secs_soc_7 + secs_eco_1 + secs_eco_2 + secs_eco_3 +
                secs_eco_4 + secs_eco_5
conservative ~ beta * family"

# Estimate the model in lavaan
results_dk <- sem(model = model_dk, data = DK)
results_us <- sem(model = model_us, data = US)

# Bind the models into a list
results <- list(results_nl, results_dk, results_us)
# Test the hypothesis that the effect size labeled 'beta' is positive
pbf(results, hypothesis = "beta > 0", standardize = TRUE)
```
```{r, eval = TRUE, echo = FALSE}
# Specify the models for DK and US
model_dk <- "
family =~ fam_1 + fam_2 + fam_3
conservative =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
conservative ~ beta * family"
model_us <- "
family =~ fam_1 + fam_2 + fam_3
conservative =~ secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
                secs_soc_6 + secs_soc_7 + secs_eco_1 + secs_eco_2 + secs_eco_3 +
                secs_eco_4 + secs_eco_5
conservative ~ beta * family"

# Estimate the model in lavaan
results_dk <- sem(model = model_dk, data = DK)
results_us <- sem(model = model_us, data = US)

# Bind the models into a list
results <- list(results_nl, results_dk, results_us)
# Test the hypothesis that the effect size labeled 'beta' is positive
bf_tot <- pbf(results, hypothesis = "beta > 0", standardize = TRUE)
bf_tot
```

As the BF in all three samples is positive, the resulting PBF is very large, $PBF `report(bf_tot[1,1])`$.
We can thus conclude that the central hypothesis receives overwhelming evidence across samples.
<!-- In the first use-case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`. -->
<!-- `pbf()` supports lists of all objects that `bain()` also supports. All supported types are specified in bain vignette *REF*. One exception is the use of named vectors, which will be covered later. -->
<!-- ```{r} -->
<!-- lm_sample1 <- lm(postnumb~site-1, sample1) -->
<!-- lm_sample2 <- lm(postnumb~site-1, sample2) -->
<!-- lm_sample3 <- lm(postnumb~site-1, sample3) -->

<!-- lm_allsamples <- list(lm_sample1, lm_sample2, lm_sample3) -->

<!-- pbf(lm_allsamples, hypotheses) -->
<!-- ``` -->

<!-- The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually. The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low. When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually. For sample 3 the specified hypotheses are even more likely than their complement. This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation. If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation. This leads to a more extreme combined evaluation than any individual evaluation. Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true. -->

<!-- Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.  -->
<!-- To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the pbf calculated with `pbf()`.  -->
<!-- In this example, data to evaluate $H_i$ is available for three countries. -->

<!-- In the first use-case of `pbf()`, an `lm()` is ran for each country.  -->
<!-- All `lm` objects are stored in a list which is passed to `pbf()`. -->
<!-- `pbf()` supports lists of all objects that `bain()` also supports.  -->
<!-- All supported types are specified in the `bain` vignette.  -->
<!-- One exception is the use of named vectors, which will be covered later. -->
<!-- ```{r} -->
<!-- fit_US <- lm(soc ~ grp + her, data = USa) -->
<!-- fit_DK <- lm(soc ~ grp + her, data = DKa) -->
<!-- fit_NL <- lm(soc ~ grp + her, data = NLa) -->

<!-- all_fits <- list(fit_US, fit_DK, fit_NL) -->

<!-- pbf(all_fits, hypotheses) -->
<!-- ``` -->

### Using bain objects

The `pbf()` function also accepts multiple `bain` objects.
This makes it possible to, for example, evaluate different sets of hypotheses on different data sets before using the resulting `bain` objects to aggregate the evidence for all common hypotheses across datasets.
The example below illustrates this use case.
As before, all analyses share one hypotheses in common ($H_i: \beta_{fam} > 0$),
but the Dutch sample now contains a sample-specific hypothesis regarding the effect of group morality, namely that $\beta_{grp} < 0$.
The `pbf()` function is called on a list of `bain` objects.
Note that, in this case, `pbf()` does not require an argument `hypothesis`, as the hypotheses are contained in the `bain` objects.

```{r, echo = TRUE, eval = FALSE}
# Add the additional predictor to the model, label the effect beta2
model_nl <- paste(model_nl, "group =~ grp_1 + grp_2 + grp_3
                             conservative ~ beta2 * group", sep = "\n")

# Estimate the model in lavaan
results_nl <- sem(model = model_nl, data = NL)

# Obtain BF for each sample; note that the Dutch sample has two hypotheses
bf_nl <- bain(results_nl, hypothesis = "beta > 0;beta2 < 0", standardize = TRUE)
bf_dk <- bain(results_dk, "beta > 0")
bf_us <- bain(results_us, "beta > 0")

# Bind bain objects into a list
bfs <- list(bf_nl, bf_dk, bf_us)

# Call pbf on that list
pbf(bfs)
```

As can be seen, the results are equivalent to the results in the previous example.
The sample-specific hypothesis has been left out, and common hypotheses are retained and aggregated.
If there are no common hypotheses across all objects, `pbf()` throws an error.

```{r, echo = FALSE, eval = FALSE}
bains_different_hyps <- list(
  bain(fit_US, "grp = 0"),
  bain(fit_DK, "grp > .1"),
  bain(fit_NL, "grp < .1")
)

tryCatch({
  pbf(bains_different_hyps)},
  error = function(e){message(e)}
      )

```
<!--The function returns an error stating that the samples have no hypothesis in common and thus the PBF cannot be calculated.-->

### Using sufficient statistics

A third use-case occurs when the raw data from different samples is not available.
This may happen, for example, when aggregating findings from the published literature (similar to meta-analysis).
In this case, one can use the default interface of `bain`, as explained in [@hoijtink2019tutorial].
This function requires four arguments: A named vector of parameter estimates, their asymptotic covariance matrix, the original sample size, and the number of within-group and between-group parameters.
Note that, when analyzing a single parameter per sample, the standard error is sufficient to construct the asymptotic covariance matrix.
Thus, this method is suitable for data that have been prepared for meta-analysis.
Importantly, unlike meta-analysis, the present method is suitable for conceptual replications.
It does not require uniform effect size measures across studies.
The example below illustrates how to aggregate evidence in favor of one hypotheses across three studies that each used different methods.
<!-- CJ: What I'd like to do here is show how to aggregate a t-test, a regression coefficient and a correlation coefficient -->

The present use case effectively tests the following hypothesis:
*There is a positive association between family morality and political conservatism*.
This conceptual hypothesis is tested differently in the three samples, resulting in three different types of statistics and distinct sample-specific hypotheses:

1. A t-test was performed using the NL data; using Cohen's D gives $H_i^{NL}: \delta_{conservative > liberal} > 0$
1. A bivariate regression coefficient for `grp` was calculated using the DK data, giving $H_i^{DK}: b_{fam} > 0$
1. A correlation coefficient was calculated using the US data, giving $H_i^{US}: \rho_{fam,con} > 0$

Note that we intentionally manipulate the data to fit these types of analyses;
for example, we compute mean scale scores and dichotomize a continuous variable to conduct a t-test.
We do not advocate these practices for applied research.

First we obtain the relevant parameter estimates and their sampling variances,
which allows us to test the specific hypotheses in `bain`:

```{r eval = TRUE, echo = TRUE}
# Create mean scale scores
NL <- data.frame(
  family = rowMeans(NL[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(NL[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
                               "sepa_eco_5")]))
DK <- data.frame(
  family = rowMeans(DK[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(DK[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
                               "sepa_eco_5")]))

US <- data.frame(
  family = rowMeans(US[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(US[c("secs_soc_1", "secs_soc_2", "secs_soc_3",
                               "secs_soc_4", "secs_soc_5", "secs_soc_6",
                               "secs_soc_7", "secs_eco_1", "secs_eco_2",
                               "secs_eco_3", "secs_eco_4", "secs_eco_5")]))

# NL: Conduct t-test using Cohen's D
NL$group <- cut(NL$conservative, breaks = 2, labels = c("liberal", "conservative"))
sample_sizes <- table(NL$group)
sds <- tapply(NL$family, NL$group, sd)
pooled_sd <- sqrt(sum((sample_sizes - 1) * sds) / (sum(sample_sizes) - 2))
NL_est <- diff(tapply(NL$family, NL$group, mean)) / pooled_sd
NL_var <- (sum(sample_sizes) / prod(sample_sizes)) + (NL_est^2 / (2*sum(sample_sizes)))

# DK: Conduct bivariate regression
DK_fit <- lm(conservative ~ family, data = DK)
DK_est <- coef(DK_fit)["family"]
DK_var <- vcov(DK_fit)["family", "family"]

# US: Correlation coefficient
US_est <- cor(US)[1, 2]
US_var <- (1 - US_est^2)^2 / (nrow(US) - 1)

# Name the estimates so hypotheses will be the same
names(NL_est) <- names(DK_est) <- names(US_est) <- "parameter"
```

Then, we use `bain.default()` to evaluate the central hypothesis on each parameter estimate. 
The pbf can then be calculated as in the previous use case.

```{r}
# Use bain.default() to obtain BF for the central hypothesis 
NL_bain <- bain(x = NL_est, 
                Sigma = matrix(NL_var, 1, 1),
                n = nrow(NL),
                hypothesis = "parameter > 0",
                joint_parameters = 1)
DK_bain <- bain(x = DK_est,
                Sigma = matrix(DK_var, 1, 1),
                n = nrow(DK),
                hypothesis = "parameter > 0",
                joint_parameters = 1)
US_bain <- bain(x = US_est,
                Sigma = matrix(US_var, 1, 1),
                n = nrow(US),
                hypothesis = "parameter > 0",
                joint_parameters = 1)

# Aggregate evidence using pbf()
pbf(list(US_bain, DK_bain, NL_bain))
```

The results suggest substantial evidence in favor of the hypothesis that there is a positive association between family morality and political conservatism.
Although each study used a different method to assess this hypothesis,
their evidence can be synthesized using the pbf.

<!--
```{r, eval = FALSE, echo = FALSE}
get_estimates <- function(sample){
  number_sites <- length(levels(sesamesim$site))

  sample_size <- nrow(sample)

  site_estimates <- sapply(1:number_sites, function(i){
    rbind(
      mean(sample[sample['site'] == i, "postnumb"]),
      sd(sample[sample['site'] == i, "postnumb"])
      )
    })

  est_names <- sprintf("site%d", 1:number_sites)
  site_means <- site_estimates[1,]
  names(site_means) <- est_names

  covmat <- diag((site_estimates[2,] / sqrt(sample_size))^2)
  dimnames(covmat) <- list(est_names, est_names)
  return(list(means = site_means, covmat = covmat, sample_size = sample_size))
}

ests_sample1 <- get_estimates(sample1)
ests_sample2 <- get_estimates(sample2)
ests_sample3 <- get_estimates(sample3)

```

Then, each object is converted to a `bain` object with these estimates
```{r}
bain_sample1_ests <- bain(x = ests_sample1[['means']], 
                          Sigma = ests_sample1[['covmat']], 
                          n = ests_sample1[['sample_size']], 
                          hypothesis = hypotheses)

bain_sample2_ests <- bain(x = ests_sample2[['means']], 
                          Sigma = ests_sample2[['covmat']], 
                          n = ests_sample2[['sample_size']], 
                          hypothesis = hypotheses)

bain_sample3_ests <- bain(x = ests_sample3[['means']], 
                          Sigma = ests_sample3[['covmat']], 
                          n = ests_sample3[['sample_size']], 
                          hypothesis = hypotheses)

bain_allsamples_ests = list(bain_sample1_ests, bain_sample2_ests, bain_sample3_ests)
```

And finally, on these objects, `pbf()` can be called.
```{r}
pbf(bain_allsamples_ests)
```
-->
<!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use-case, or is something off in the script? -->
<!-- CJ: We should ask Herbert about this, but this example won't make it into this paper. If you want, you can send Herbert a reproducible example of the 'problem'!-->

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup