### 2023-04-25 manuscript.Rmd
Rebecca's edits

---
title             : "Aggregating evidence from conceptual replication studies using the product Bayes factor"
shorttitle        : "PRODUCT BAYES FACTOR"

author:
  - name: "Caspar J. Van Lissa"
    affiliation: "1,2"
    corresponding: yes
    address: "Padualaan 14, 3584CH Utrecht, The Netherlands"
    email: "c.j.vanlissa@uu.nl"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Formal Analysis
      - Funding acquisition
      - Methodology
      - Project administration
      - Software
      - Supervision
      - Writing – original draft
      - Writing – review & editing
  - name: "Eli-Boaz Clapper"
    affiliation: "2"
    role:
      - Formal Analysis
      - Writing – original draft
      - Writing – review & editing
  - name: "Rebecca Kuiper"
    affiliation: "2"
    role:
      - Conceptualization
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Tilburg University, dept. Methodology & Statistics"
  - id            : "2"
    institution   : "Utrecht University, dept. Methodology & Statistics"

authornote: |
  This is a preprint paper, generated from Git Commit # `r substr(gert::git_commit_id(),1,7)`. This work was funded by a NWO Veni Grant (NWO Grant Number VI.Veni.191G.090), awarded to the lead author.

abstract: |
  The product Bayes factor (PBF) can synthesize evidence for an informative hypothesis across heterogeneous replication studies. It is particularly useful when the number of studies is relatively low and conventional assumptions about between-studies heterogeneity are likely violated. The present paper introduces a user-friendly implementation of the PBF in the bain R-package, and demonstrates its use in several tutorial examples. The method was validated in a simulation study that manipulated sample size, number of replication samples, and reliability. PBF had a higher overall accuracy when benchmarked against other evidence synthesis methods, including random-effects meta-analysis (RMA). This was primarily due to PBF’s greater sensitivity in detecting a true effect. However, PBF had relatively lower specificity. The PBF showed increasing sensitivity and specificity with increasing sample size. With an increasing number of samples, lower sensitivity was traded for greater specificity. Although PBF's overall performance was less susceptible to reliability than the other algorithms, this masked a trade-off between reliability and specificity. PBF thus appears to be a promising method for meta-analysis of heterogeneous conceptual replication studies. Nonetheless, users should be aware of its lower specificity, and the fact that it takes a Bayesian approach to inference and addresses a qualitatively different research question than other evidence synthesis methods.

keywords          : "bayes factor, evidence synthesis, bayesian, meta-analysis"
wordcount         : "5356"

bibliography      : ["product_bayes.bib", "references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
knit: worcs::cite_essential
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
run_everything <- TRUE
library("papaja")
library(tidySEM)
library(kableExtra)
library(pwr)
library(bain)
r_refs("r-references.bib")
source("UI function/pbf.R")
out <- readRDS("out.RData")
resultaction = "markup"
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning = FALSE)
```

Recent years have seen a crisis of confidence over the reliability of published results in psychology, and science more broadly [@brembs2018prestigious].
Replication research has emerged as one potential way to address this crisis and derive knowledge that will stand the test of time [see @lavelleWhenCrisisBecomes2021]. 
In step with this interest in replication research,
research synthesis methods have become increasingly popular.
These methods aggregate research findings,
and thus enable drawing overarching conclusions across multiple (replication) studies.
This paper addresses Bayesian evidence synthesis,
a research synthesis method that aggregates evidence for an informative hypothesis, quantified by the Bayes factor,
across multiple studies.
This method has the potential to provide a more comprehensive and accurate picture of the state of the literature, and to identify areas of consensus and disagreement among studies.
We describe the method in detail,
benchmark its performance against other commonly used research synthesis methods,
and demonstrate its application through a tutorial example analysis.
To facilitate uptake of the method by applied researchers, we introduce an implementation of the method in the `bain` R-package.
This implementation enables the use of Bayesian evidence synthesis for many commonly used statistical analyses in R.

A key challenge in quantitative research synthesis is dealing with between-studies heterogeneity [@higginsReevaluationRandomeffectsMetaanalysis2009].
When studies examine the same research question in different laboratories, use idiosyncratic methods, and sample from distinct populations, these between-study differences can introduce heterogeneity in findings.
The most common quantitative research synthesis method is meta-analysis,
in which results of different studies are aggregated to estimate an aggregate effect size [@borensteinIntroductionMetaAnalysis2009].
In meta-analysis, heterogeneity can be accounted for in four ways [see @vanlissaSmallSampleMetaanalyses2020].
First, if studies are exact replications,
one may assume that no heterogeneity in the outcome exists and a fixed-effect meta-analysis can be conducted to estimate the common population effect. 
Second, when heterogeneity between studies can be assumed to be random,
random-effects meta-analysis can be used to estimate the mean of a distribution of population effects. 
Third, when there are a few systematic differences between studies, these can be accounted for using meta-regression.
Finally, when there are many potential variables that cause systematic differences and it is not known beforehand which are relevant, exploratory techniques like random forest meta-analysis and penalized meta-regression can be used to identify relevant moderators [@vanlissaSelectRelevantModerators2021].
However, accounting for moderators requires a relatively high number of observations per moderator,
which may not be available.
Each of these approaches thus makes different assumptions about the nature of heterogeneity [see "Models for meta-analysis" in @vanlissaSmallSampleMetaanalyses2020].

A crucial shortcoming of existing research synthesis methods is that the aforementioned assumptions
may not be tenable when meta-analyzing studies that investigate the same informative hypothesis,
but are otherwise very heterogeneous.
The situation may arise where each study is uniquely identified by a combination of linearly dependent moderators.
In this case, it is no longer possible to synthesize *effect sizes* while accounting for heterogeneity using statistical methods.
<!-- even machine learning based methods cannot identify relevant moderators. -->
It is still possible, however, to quantify the support these studies provide for the underlying informative hypothesis.
To this end, Bayesian evidence synthesis (BES) aggregates the evidence for a theoretical relationship across studies,
without imposing assumptions about heterogeneity [@kuiperCombiningStatisticalEvidence2013].
Although this assumption is not necessary,
for the sake of simplicity we assume that this theoretical relationship is evaluated via informative hypothesis $H_i$ in all studies.

The amount of evidence for a hypothesis can be expressed as a Bayes factor, or BF.
The BF can be interpreted as the ratio of evidence for one hypothesis relative to another hypothesis.
<!-- , so $BF_{12} = \frac{BF_{1u}}{BF_{2u}}$. -->
<!-- Note that $BF_{12}$ is itself a ratio of two Bayes factors with a common denominator, which renders them comparable [see @guApproximatedAdjustedFractional2018]. -->
Within the scope of this paper,
all Bayes factors are the ratio of evidence for an informative hypothesis $H_i$ relative to its complement $H_{!i}$ [see @guApproximatedAdjustedFractional2018].
<!-- This Bayes factor , $BF_{c} = \frac{H_i}{H_!i}$.  -->
The subscript $_!$ represents the negation operator;
in other words, $H_{!i}$ means "not $H_i$".
This Bayes factor, which we will refer to as $BF_c$, represents the ratio of evidence for $H_i$ divided by evidence against it.
A value of $BF_c = 10$ means that the data provide ten times more support for the hypothesis than against it.

When multiple studies each provide evidence for $H_i$ in the form of complement Bayes factors,
these Bayes factors can be synthesized across studies by taking their product [@kuiperCombiningStatisticalEvidence2013].
The resulting product Bayes factor (PBF) summarizes the total evidence for the hypothesis.
The only assumption of the PBF is that all study-specific hypotheses provide evidence about the same underlying theoretical relationship.
Note that other approaches to BES exist;
for instance, it is possible to use the posterior of one study as the prior for a replication study, and thus accumulate evidence across studies [see @heckReviewApplicationsBayes2022].
Such applications are out of scope of the present paper,
which addresses the PBF approach to BES.

Although meta-analysis and BES are both research synthesis methods, they answer different research questions.
Meta-analysis estimates the point estimate or distribution of a population effect size. 
It pools estimates of this effect size across multiple studies to obtain an overall estimate of the effect size.
It thus answers questions like:
Given certain assumptions about between-studies heterogeneity, what is the average population effect size?
BES, on the other hand, aggregates evidence for an informative hypothesis across multiple studies.
It thus answers the question: Do all these studies support the hypothesis of interest?
Both methods are appropriate for different research questions, and provide complementary information.

This paper introduces the first implementation of BES in user-friendly free open source software (FOSS).
A function `pbf()` was contributed to the `bain` R-package for Bayesian informative hypothesis evaluation, version `0.2.9`.
<!-- `r as.character(packageVersion("bain"))`. -->
This paper presents a simulation study to validate the method and benchmark it against alternative evidence synthesis methods,
and illustrates several use cases through reproducible examples.

# Simulation study

<!-- Rebecca: -->
<!-- ms meer over het model vertellen. -->
<!-- Is het voor de lezer duidelijk waar ineens deze groepen vandaan komen? Zijn dat je aantal studies die je gaat samennemen? -->
<!-- Dus bijv: -->
<!-- In de populatie zijn er twee variabelen, die cooreleren met de waarde rho. We hebben k studies die elk een sample van n hebben getrokken/verzameld. Dan vertelen wat de verschillende algrotithms doen met de verschillende samples (meeste als niet alle: In elke sample wordt de sample correlatie bepaald; dan estimates samenemen en null hypo test of in each sample bewijs voor hypo en dat samennemen). Dan iets over de verschillende waardes die je bekijkt en dat elke combi een simulatie conditie is. En dan iets als: Dit doen we 1000 keer per set van condities.  -->
<!-- En dan nog iets over reliability ook. -->
The present simulation study set out to validate the PBF algorithm and benchmark it against other evidence synthesis methods.
We simulated a scenario where an informative hypothesis about a correlation between two variables was measured across several independent samples,
and the resulting evidence was synthesized across samples using multiple methods.
<!-- Rebecca: -->
<!-- waar is > 3 op gebaseerd? -->
<!-- Kass&Raftery? Als ja, bedenk wel dat dat om equaility restricties gaat (bij mijn weten). Je kan het natuurlijk gebruiken, maar plaats dan wel de kanttekening, maar is een uitgangspunt natuurlijk.  -->
<!-- Of vertel over betting odds... -->
<!-- Maar het blijft een dichitome beslissing zo.... -->
<!-- Laat je het gemiddelde PBF met range ofzo zien.... zo niet, dan zou dat toevoegen. Dan zie je meer dan een dichitome beslissing. -->
<!-- En ms kan je ook ene keer 1 en 2 gebruiken, gewoon om te zien wat er gebeurt? -->
<!-- Rebecca: -->
<!-- Now, you will say that selecting > .1 with rho .1 is false, but it is not false. It is on the border, so Hi is true.  -->
<!-- In practice it may be true half of the time, because of sampling, but still. -->
<!-- This should be reflected by BF is approx 1. -->
<!-- In your simulation, you use PBF > 3, so this may be hard to reach when the truth is on the border, but that is due to the subjective choice of 3... -->
<!-- So, I already thought that using a smaller population rho value would be of added value, but now I believe you should really include this. -->
<!-- I do see that this would end up in multiple results (once using .1 and 0.2, once (say) 0 or 0.05 and .2; but also when using other cut-off values for PBF). -->
The informative hypothesis, set to be equal across studies,
was $H_i: \rho > .1$.
To examine the performance of the different evidence synthesis methods in a range of scenarios,
several design factors were manipulated.
First was the presence or absence of a true population effect.
Given the informative hypothesis of $H_i: \rho > .1$,
the presence of a true population effect was defined as $\rho = .2$ and
a null effect was defined as $\rho = .1$.
The second design factor was the number of observations per sample $n \in (50, 200, 500, 800)$.
These values were chosen because they correspond to a statistical
<!-- pwr.r.test(c(50, 200, 500, 800), .1) -->
power to reject a false null hypothesis of $\beta \in (.10, .30, .60, .80)$ power, respectively,
assuming $\alpha = .05$ and a known effect size of $\rho = .1$ [@cohenStatisticalPowerAnalysis1988].
Third, we manipulated the number of independent samples (or: replication studies), $k \in (2, 3, 10)$.
Fourth, the reliability of the two correlated variables was varied between $\alpha \in (0.6, 0.8, 1.0)$ to range from questionable to perfect reliability [@nunnallyPsychometricTheory2017].
Questionable reliability is the lowest level considered to be acceptable in social scientific research, and perfect reliability is what is assumed when analyzing correlations between observed items or scale scores.
For all unique combinations of these design factors,
the simulation was repeated 1000 times.

## Algorithms

The main algorithm of interest was the PBF.
As a decision criterion to conclude that $H_i$ was supported over its complement,
we used $PBF > 3$ - a conventional threshold for inference using Bayes factors [@jeffreys1998theory].

As a benchmark for comparison, we included several other algorithms that might feasibly be used by researchers who intend to examine whether a hypothesis is true across several independent samples.
The first was *vote counting*: counting the number of significant effects.
Although this method is still in use for aggregating conceptual replications,
it is considered bad practice.
Three disadvantages are that vote counting disregards sample size,
reduces statistical power,
and does not quantify the strength of the evidence [@hedgesVotecountingMethodsResearch1980].
Our vote counting algorithm summed the number of one-sided z-tests of a null hypothesis corresponding to the informative hypothesis,
so $H_0: \rho = .1$ and $H_a: \rho > .1$, which corresponds to $H_i$.
The decision criterion was that the hypothesis was supported in the majority of samples.
Thus, for example, if $H_0$ was rejected in three out of five samples, our vote counting algorithm would find overall support for $H_a$ (and, by extension, $H_i$).

The third algorithm was *random-effects meta-analysis* (RMA),
which is the current gold standard for evidence synthesis [@viechtbauerConductingMetaanalysesMetafor2010].
For this algorithm, the null-hypothesis was rejected if the lower bound of a 90% confidence interval for the overall effect size excluded $H_i$.
Note that this corresponds to a test at $\alpha = .05$,
because all effects in the simulation are directional.

The fourth algorithm was *individual participant data* (IPD) meta-analysis [@rileyMetaanalysisIndividualParticipant2010].
Like classic meta-analysis (RMA), IPD is a multilevel model, clustered by sample.
IPD uses the raw data, which makes it possible to estimate variance at the first level.
By contrast, RMA treats the first-level variance as known.
Note that the PBF can be estimated using either sufficient statistics (as in meta-analysis) or using raw data (as in IPD).
With this in mind, it is informative to benchmark it against both of these methods.
IPD was also evaluated using the lower bound of a 90% confidence interval for the overall effect size.
<!-- Rebecca: -->
<!-- zou je dit vooraf moeten zeggen? -->
<!-- Dan zeggen dat ILP dit ook kan en dan zeggen dus dat ook als vergelijking meenemen? -->
<!-- Ms zelfs in intro dit al noemoen, dan zeggen dat er twee meta-an equivalenten zijn, waarin je null hypo testing kan doen. En we dat gaan meenemen om te vergelijken. -->

## Performance indicators

For each algorithm, inferential decisions made using the criteria described above were compared to the population status of the hypothesis (true or false).
The resulting confusion matrix gives the number of decisions that were true positives (TP), true negatives (TN), false positives (FP), and false negatives (FN).
These quantities were summarized as sensitivity, $\frac{TP}{TP + FN}$,
the ability to detect an effect given that it was indeed true in the population,
<!-- i think sensitivity is the ability to detect an effect, given that the effect is truly there. I found that its formula is usually presented as TP/(TP+FN)-->
<!--CJ: This is all the same, see https://en.wikipedia.org/wiki/Sensitivity_and_specificity -->
and specificity, $\frac{TN}{TN+FP}$,
the ability to correctly conclude that the informative hypothesis is not supported,
given that it was indeed false in the population. <!-- and specificity is the ability to not acknowledge an effect, given that there indeed is none-->
The overall performance was captured by the accuracy, which represents the total proportion of correct (true positive and true negative) decisions, $\frac{TP + TN}{TP+TN+FP+FN}$.

```{r, echo = F}
library(data.table) # load in libraries
library(DT)
library(ggplot2)
# rename <- c("prodbf" = "PBF", "gpbf" = "gPBF", "tbf" = "TBF", "ipd" = "IPD", "rma" = "RMA", "allsig" = "VC")

```

# Results

```{r tabconf, echo = F, message=F}
tab <- read.csv("confusion.csv", stringsAsFactors = FALSE)
names(tab)[1] <- "Metric"
tab$Metric <- gsub("^.+_", "", tab$Metric)
tab <- tab[!tab$Metric == "lr", !grepl("iu", names(tab))]
names(tab) <- gsub("_ic", "", names(tab))
# names(tab)[match(names(rename), names(tab))] <- rename
tab <- tab[tab$Metric %in% c("sensitivity", "specificity", "accuracy"),]
rownames(tab) <- NULL
knitr::kable(tab, digits = 2, caption = "Marginal confusion matrix metrics.")
```
We examined overall model performance across conditions.
PBF had a higher overall accuracy than other algorithms
followed by IPD, then RMA, and finally VC,
see Table \@ref(tab:tabconf).
This higher accuracy was primarily driven by PBF's greater sensitivity to detect a true effect compared to other algorithms.
However, PBF had a lower specificity compared to all other algorithms.
This suggests that the PBF trades a loss of specificity for increased sensitivity.

## Effect of simulation conditions

```{r tabeffect, echo = FALSE}
tab <- read.csv("effect_of_conditions.csv", stringsAsFactors = FALSE)
tab <- tab[, !grepl("(gpbf|tbf)", names(tab))]
tab <- tab[, grepl("^\\w+$", names(tab)) | (grepl("\\.{2}", names(tab)) & grepl("PBF", names(tab)))]
names(tab)[grepl("\\.{2}", names(tab))] <- paste0("vs ", gsub("(\\.|PBF|vs)", "", names(tab)[grepl("\\.{2}", names(tab))]))
knitr::kable(tab, digits = 2, caption = "Partial eta squared of the effect of each design factor on accuracy for each algorithm and for the difference between PBF and all other algorithms (e.g., vs RMA).")
```
We used ANOVAs to examine the effect of simulation conditions on overall accuracy.
The differences between algorithms were analyzed in analyses that included two-way interactions between design factors and algorithm.
As the sample size was very large,
significance tests were uninformative.
We thus focused on interpreting the effect size of the effects of design factors.
The performance of PBF was most impacted by sample size $n$, followed by the number of groups $k$, and reliability.
The differences in the effects of sample size and number of groups were relatively small between PBF and the two best practice algorithms, RMA and IPD -
but substantial between PBF and the suboptimal VC algorithm.
The reverse pattern occurred for reliability, however:
it showed a substantial difference in effect between PBF and the two best practice algorithms only.

### Effect of sample size

```{r}
res <- readRDS("confusion_by_cond.RData")

df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("n", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("n", "Outcome", "alg")]
library(ggplot2)
p <- ggplot(df_plot, aes(x = n, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  scale_linetype(guide = guide_legend(reverse = TRUE) )+
  theme_bw() +
  labs(y = "Mean performance", x = "Sample size")+
  labs(shape = "Algorithm")
ggsave("sample_size.pdf", p, device = "pdf", width = 140, height = 100, units = "mm")
```
```{r fign, fig.cap="Mean performance by sample size"}
knitr::include_graphics("sample_size.pdf")
```

Figure \@ref(fig:fign) indicates that for PBF, both sensitivity and specificity increased with sample size.
The other algorithms showed only increasing sensitivity; 
specificity was limited by a ceiling effect.
This difference explains the effect of sample size on the difference between algorithms (see Table \@ref(tab:tabeffect)).

### Effect of number of samples

```{r}
df_plot <- data.table::melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("k", "Outcome", "alg")]
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("k", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
p <- ggplot(df_plot, aes(x = k, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  scale_linetype(guide = guide_legend(reverse = TRUE) )+
  theme_bw() +
  labs(y = "Mean performance", x = "Number of groups")+
  labs(shape = "Algorithm")
ggsave("groups.pdf", p, device = "pdf", width = 140, height = 100, units = "mm")
```
```{r figk, fig.cap="Mean performance by number of groups"}
knitr::include_graphics("groups.pdf")
```

Figure \@ref(fig:figk) indicates that, for PBF at higher levels of $k$,
lower sensitivity was traded for greater specificity.
The other algorithms did not show this pattern, as their specificity was at a ceiling.
This difference in pattern of effects explains why number of samples had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Only VC showed decreasing sensitivity with an increasing number of groups;
this is because the probability of obtaining false negatives increases with the number of groups [as also noted by @hedgesVotecountingMethodsResearch1980].

### Effect of reliability

<!--EB: I think reliability plot is a bit confusing (The one that is saved under ./manuscript_files/figure-latex/unnamed-chunk-3-1.png). That is because the x-axis is labeled 'reliability', but the values are the values of the 'errorsd' condition. This makes that the reliability decreases, rather than increases as you move away from the origin on the x-axis -->
```{r}
df_plot <- melt(res, measure.vars = c("sensitivity", "specificity"),
             variable.name = "Outcome", value.name = "val")
df_plot$reliability[df_plot$reliability == 0] <- 1
df_plot$reliability[df_plot$reliability == .5] <- .8
df_plot$reliability[df_plot$reliability == .81] <- .6

# df_plot <- df_plot[ ,list(mean=mean(val)), by=c("reliability", "Outcome", "alg")]
df_plot <- as.data.frame(df_plot)
df_plot <- aggregate(df_plot$val, df_plot[c("reliability", "Outcome", "alg")], FUN=mean)
names(df_plot)[ncol(df_plot)] <- "mean"
p <- ggplot(df_plot, aes(x = reliability, y = mean, group = interaction(Outcome, alg), shape = alg, linetype = Outcome)) +
  geom_point() +
  geom_line() +
  scale_linetype(guide = guide_legend(reverse = TRUE) )+
  theme_bw() +
  labs(y = "Mean performance", x = "Reliability")+
  labs(shape = "Algorithm")
ggsave("reliability.pdf", p, device = "pdf", width = 140, height = 100, units = "mm")
```
```{r figrel, fig.cap="Mean performance by reliability"}
knitr::include_graphics("reliability.pdf")
```

Figure \@ref(fig:figrel) indicates that the PBF traded sensitivity for specificity.
At low levels of reliability, specificity exceeded sensitivity;
at high levels of reliability, this pattern was reversed.
The other algorithms did not show this pattern, as their specificity was at a ceiling.
Their sensitivity increased with higher reliability, however,
and therefore, so did their overall performance.
This difference in pattern of effects explains why reliability had a moderate effect on the difference between algorithms (see Table \@ref(tab:tabeffect)).
Note that, whereas the overall accuracy of the PBF was found to be less susceptible to reliability as compared to other algorithms,
this finding masked the trade-off between reliability and specificity.

# Discussion

This simulation study examined the performance of the product Bayes factor (PBF) as compared against three other common methods for evidence synthesis (IPD, RMA, and VC).
The results showed that PBF had a higher overall accuracy than the other algorithms,
primarily due to its greater sensitivity to detect a true effect.
PBF had lower specificity, however,
suggesting that it trades specificity for increased sensitivity.
<!-- PBF's performance was most impacted by sample size, both in terms of sensitivity and specificity. The other algorithms showed less improvement in accuracy with increasing sample size due to a ceiling effect for specificity. -->
<!-- The number of groups also had a moderate effect on PBF's performance, -->
<!-- increasing sensitivity while specificity remained approximately stable. -->
<!-- Reliability had a moderate effect on the performance of all algorithms, with overall performance increasing as reliability increased. -->
<!-- However, the effect of reliability on PBF was different than the other algorithms, as PBF traded increasing sensitivity for decreasing specificity as reliability increased, -->
<!-- whereas the other algorithms showed only an increase in sensitivity and approximately stable specificity. -->
PBF's performance was most impacted by sample size,
followed by the number of groups and reliability.
Even at the smallest sample size of $n = 20$,
PBF had a superior sensitivity to the other algorithms.
This suggests that PBF is more suitable than other methods as a small sample solution.
When the number of samples increased,
most algorithms showed approximately stable specificity
and increasing sensitivity.
A notable exception was vote counting;
its sensitivity decreased with an increasing number of samples,
as has been previously documented [@hedgesVotecountingMethodsResearch1980].
With increasing reliability,
PBF showed a substantially different pattern of results than the two other best practice algorithms (RMA and IPD).
Although the overall effect of reliability on overall accuracy was smaller for PBF than for other algorithms,
this masks a trade-off between reliability and specificity.
Whereas most algorithms showed near-stable specificity and increasing sensitivity,
PBF showed a clear trade-off between decreasing specificity and increasing sensitivity.
This implies that the PBF is more conservative - less likely to detect an effect - in the presence of increasing measurement error.

These results have important implications for applied evidence synthesis.
For example, the finding that PBF had a higher overall accuracy due to greater sensitivity suggests that it may be a better choice than the other algorithms,
particularly when detecting true effects has high priority.
However, researchers should be aware that this increased sensitivity comes at a loss of specificity,
which incurs a greater risk of false positive results.
If specificity is a higher priority,
other algorithms such as IPD or RMA may thus be more appropriate.

The present study also has some limitations.
First, the simulation study makes specific assumptions that may not generalize to all real-world applications.
A second important caveat is that most of the algorithms did not reach
a level of sensitivity that is considered acceptable from a perspective of statistical power [i.e., greater than .80, @cohen2013statistical].
PBF performed notably better than other algorithms,
but its sensitivity still fell below .80 in many conditions.
Low power increases the risk of false negatives,
or failing to detect a true effect.
One reason power was low is that,
in conditions where a true effect was present,
its value only exceeded the boundary value of the informative hypothesis by .1 points.
Such small effects are hard to detect.
All algorithms will likely perform better when the true effect is larger.
The low sensitivity of all algorithms highlights the importance of reticence when interpreting evidence syntheses of studies with small samples and small effect sizes.
It may be prudent to avoid generalizing such results to the population,
and instead consider them as merely descriptive of the published research.
Additionally, sensitivity analyses can be used to assess the robustness of the results to different modeling assumptions and methods.

A third limitation is that the evidence synthesis methods compared here represent different approaches to inference and answer different research questions.
<!-- Their inferential properties are thus not directly comparable. -->
<!-- PBF assesses whether an informative hypothesis is supported across all replication studies; -->
<!-- RMA estimates an average effect size from summary statistics; -->
<!-- IPD estimates an effect size from individual data, accounting for clustering within studies; -->
<!-- and VC counts the number of times a hypothesis was supported. -->
Since each of these methods is optimized for a different purpose,
the present study should not be considered as a comprehensive assessment of their strengths and weaknesses.
<!-- For example, PBF takes a Bayesian approach to inference, and a PBF of 10 in our study means that the (subjective) probability that our informative hypothesis is true is ten times larger than the probability that it is not true. -->
<!-- By contrast, IPD takes a frequentist approach to inference, -->
<!-- and a p-value of $p = .02$ means that there is a 2% probability of observing an effect at least as large as the one found, if the null-hypothesis were true. -->
We nonetheless compare them because of their similar usage in evidence synthesis.
It is up to individual researchers to choose an appropriate method,
guided by the research question and the available information.
For instance, when raw data is unavailable, IPD cannot be used,
and when parameter estimates or effect sizes are not reported, only VC can be used.

Aside from the aforementioned fact that the PBF answers a different research question than the other algorithms,
it is worth noting limitations of the interpretation of the PBF.
The PBF renders support for one specific informative hypothesis versus its complement.
If the informative hypothesis is supported, this does not necessarily mean that it is also true.
Consider the hypothetical example that the informative hypothesis that the earth is flat was supported with $BF = 3.01$.
Although the data support this hypothesis over its complement, the hypothesis is clearly wrong (the earth is spherical).
If we would have evaluated another hypothesis, e.g., the earth is shaped like an American football, it would have received much more support, e.g. $BF = 1000$, even though it is also wrong.
A high Bayes factor thus does not mean that the hypothesis is true.
Conversely, a low Bayes factor merely indicates that the informative hypothesis is not supported,
and does not provide information about the true state of affairs.
A related limitation is that our simulation study used an arbitrary - albeit conventional - threshold for inference [@jeffreys1998theory].
In applied research, it is more sensible to evaluate the weight of evidence, rather than resorting to a rule of thumb.
<!-- In sum, this simulation study suggest that PBF may be a good choice when sensitivity is of utmost importance, but researchers should be mindful of the trade-off between specificity and sensitivity. -->
<!-- The results also highlight the importance of considering sample size, number of samples, -->
<!-- and reliability when designing studies and conducting meta-analyses. Future research should examine the performance of additional algorithms and the impact of other design factors on algorithm performance. -->
<!-- The results indicated that, compared to the other algorithms, -->
<!-- PBF had a higher overall accuracy, -->
<!-- which could be attributed to PBF's relatively greater sensitivity to detect a true effect. -->
<!-- The greater sensitivity of PBF came at a loss of specificity compared to other algorithms. -->
<!-- Specifically, all other algorithms showed ceiling effects for specificity and low base rates for sensitivity. -->

<!-- The overall accuracy of the PBF was most affected by the sample size of the underlying studies. -->
<!-- This effect was partly driven by the absence of a ceiling effect of specificity for the PBF, compared to the other algorithms. -->
<!-- Increasing sample size increased sensitivity and specificity for all algorithms. -->
<!-- Only for PBF, however, did sensitivity surpass specificity at large ($\geq 200$) sample sizes. -->
<!-- Combined with the higher overall level of specificity, this suggests that PBF may be preferable when detecting true effects has high priority. -->

<!-- <!-- Small samples -->
<!-- At the smallest sample size of $n = 20$, the PBF had a better balance between sensitivity and specificity than the other algorithms. -->
<!-- This suggests that PBF is more suitable than other methods as a small sample size solution for evidence synthesis. -->

<!-- that increased with increasing sample size, number of samples, and reliability. -->
<!-- PBF, by contrast, showed lower base rates of specificity and higher base rates of sensitivity. -->

<!-- With increasing sample size, number of samples, and reliability, PBF showed a trade-off between decreasing specificity and increasing sensitivity. -->

<!-- Overall, these results suggest that PBF had relatively better overall accuracy than the other algorithms under consideration. -->
<!-- Although other algorithms had superior specificity, PBF had the highest levels of sensitivity. -->
<!-- PBF thus traded off an increased ability to correctly accept informative hypotheses in the presence of true population effects for a decreased ability to reject the hypothesis when there was no effect. -->
<!-- Users of the PBF should assess whether this trade-off is warranted. -->
<!-- If it is not, more stringent criteria for inference or more conservative methods may be preferred. -->

<!-- OTHER PBF APPROACHES -->
<!-- One possible issue with the PBF as defined in this research comes from an article of @rouder2011bayes.  -->
<!-- The article states that a meta-analytic Bayes factor is not as straightforward as taking the product of $k$ individual Bayes factors. -->
<!-- <!--EB: Their reason seems to be that Bayes factors take sample size into account, effectively reducing the ability of a Bayes factor to recognize an effect in small sample. I dont see how this invalidates the PBF. If there is an effect, then in 5 small studies, it may be true that a BF is calculated that barely support the hypothesis of an effect. However, taking the product of 5 BFs that are all 1.25 will result in a PBF of about 3.-->
<!-- Rather, they provide a way of calculating more valid a meta-analytic Bayes factor. -->
<!-- However, only under the assumption that the population effect is constant and the test statistic is from the same type of statistical model. -->
<!-- Using this form of the PBF, they show that pooling the data and calculating a single Bayes factor on the pooled data comes closer to the meta-analytic Bayes factor than taking the product of each individual study. -->
<!-- However, as stated earlier, one strength of the PBF lies in its ability to synthesize evidence for a hypothesis, rather than an estimate. -->
<!-- This means that there is no assumption of a constant population effect and/or equal types of test statistics across studies. -->
<!-- Furthermore, as the article of @rouder2011bayes states itself, pooling data is preferred if all data is independent and drawn from a common distribution.  -->
<!-- The study idiosyncrasies encountered in evidence synthesis should wary the researcher in blindly making the assumption of a common distribution. -->
<!-- A final remark is that of performance comparison between PBF and the other algorithms which has to do with the differences in hypothesis specification.  -->
<!-- For the frequentist and PBF algorithms we specified $H_0: \rho = .1$ and $H_c: \rho \leq .1$ respectively. -->
<!-- This means, hypothetically, that if a strong effect in the opposite direction was found, let's say $\rho = -.6$, the frequentist algorithms would reject $H_0$ while the PBF would accept $H_c$. -->
<!-- This implies that there should have been a higher power to detect an effect in the frequentist framework, as negative effects would reject $H_0$. -->
<!-- This is contrary to what was observed in this research, however, as the frequentist approaches had a low sensitivity. -->
<!-- Again, this has to do with the small simulated difference between the true effect and the boundary value. -->

# Tutorial

This tutorial demonstrates how to synthesize evidence for an informative hypothesis across heterogeneous replications using the Product Bayes Factor (PBF). 
We assume that users have installed the free open source statistical programming language R [@R-base].
The R-package `bain` version `0.2.9` or later is required, which can be installed by running `install.packages("bain")` in the R console.
The data used in this tutorial are included in the `bain` package,
and have been simulated based on the data presented in [@vanleeuwen2022morality].
A more detailed description of the datasets is found in [@vanleeuwen2022morality];
additionally, the dataset documentation is accessed by running `?synthetic_us`, `?synthetic_dk` or `?synthetic_nl` in the R console.
Van Leeuwen and colleagues conducted a theory-driven, preregistered study to address the research question whether political orientation and moral dispositions are associated.
Suitable data were collected in three countries: the United states of America, Denmark, and the Netherlands.
Each sample contained multiple measures of political orientation and moral dispositions.
In the original publication, the PBF was used to aggregate evidence across scales and countries to obtain an overall measure of support for the central hypothesis.
This tutorial follows the same rationale, but uses only one effect size per sample,
and varies the way this effect size is computed to illustrate the more typical use case where the same informative hypothesis has been studied in different ways in multiple studies.
We will examine the informative hypothesis that self-reported importance of family morality is positively associated with a conservative socio-political orientation.
We load the `bain` library and assign the data to three objects with convenient names:

```{r eval = TRUE, echo = TRUE}
library(bain)
NL <- synthetic_nl
DK <- synthetic_dk
US <- synthetic_us
```

<!-- skeleton
1.) explain calculation of single BF and what it means, refer to previous tutorial on the bayes factor
2.) introduce the case where there are multiple studies and we want to evaluate a single hypothesis
3.) introduce pbf(), its uses, outputs and interpretation
-->

### How to use bain

<!-- CJ: Is dit niet teveel herhaling van wat we in de inleiding zeggen?? -->
<!-- In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis.  -->
<!-- This method evaluates a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data.  -->
<!-- However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis. They may also not only be interested in whether the hypothesis is probable, but also by how much.  -->
<!-- In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis.  -->
<!-- Because the focus of this tutorial is on BES,  -->
We briefly introduce the basic use of the `bain()` function, and how to interpret its output.
We must estimate a model suitable for evaluating our informative hypothesis.
Because both scales consist of multiple items, we can use structural equation modeling (SEM) to perform latent variable regression [see @vanlissaTeacherCornerEvaluating2020]:
<!--EB: For simplicities sake, I for now create simple average scores without taking into account reverse coding. This should likely be improved with factor scores -->
<!-- CJ: Keep the code SUPER simple. No functions etc -->

```{r echo = TRUE, eval = TRUE}
# Load lavaan package for SEM
library(lavaan)

# Specify SEM-model for latent variable regression
model_nl <- "
fam =~ fam_1 + fam_2 + fam_3
con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
       sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
con ~ beta * fam"

# Estimate the model in lavaan
results_nl <- sem(model = model_nl, data = NL)
```
```{r echo = FALSE, eval = FALSE}
# Load relevant packages
library(bain)
library(lavaan)

# Specify a SEM-model with two latent variables and latent regression
model_nl <- "
fam =~ fam_1 + fam_2 + fam_3
con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
                sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
con ~ beta * fam"

# Estimate the model in lavaan
results_nl <- sem(model = model_nl, data = NL)

# Test the hypothesis that the effect size labeled 'beta' is positive
# The argument standardize = TRUE uses the standardized model parameters
bf_nl <- bain(results_nl, hypothesis = "beta > .1", standardize = TRUE)
```

The informative hypothesis in this tutorial is $H_i: \beta > .1$, where $\beta$ (beta) is the standardized regression coefficient.
Instead of a conventional null hypothesis, $H_0: \beta = 0$,
the value of $.1$ was used as a minimal effect size of interest.
The code below illustrates how to obtain a Bayes factor for this informative hypothesis,
using the output of the SEM analysis above.
We can refer to the parameter `beta` by name because we labeled it in the `lavaan` syntax;
if we had not done so, we could find the names of all model parameters by running `get_estimates(results_nl, standardize = TRUE)`.
The results indicate that the hypothesis is supported when compared to its complement.
<!-- , $BF_c `report(bf_nl$fit$BF.c[1])`$. -->
For a more in-depth tutorial on `bain()`, see @hoijtink2019tutorial, and for further guidance on the use of `bain()` for SEM, see @vanlissaTeacherCornerEvaluating2020.

```{r echo = TRUE, eval = TRUE, results = resultaction}
# Test that the effect labeled 'beta' is positive
bf_nl <- bain(results_nl, hypothesis = "beta > .1", standardize = TRUE)
bf_nl
```

<!-- Socio-political orientation and group loyalty have been aggregated to variables `soc` and `grp` respectively.  -->
<!-- Both are continuous variables with higher scores indicating either conservatism or higher group loyalty.  -->
<!-- In NHST we specify the null hypothesis that there is no association between `soc` and `grp`.  -->
<!-- One method to test the hypothesis is to calculate a standardized regression coefficient. -->

<!-- $H_0: \beta_{grp} = 0$. -->

<!-- However, we may have prior expectations that higher group loyalty is somewhat related to  higher conservatism.  -->
<!-- An informative hypothesis could be: -->

<!-- $H_i: \beta_{grp} > 0$ -->

<!-- SESAMESIM DATA: Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`). `postnumb` is a continuous variable which ranges from 0-63. `site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking. In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  site`:

$H_0: \mu_{site1} =\mu_{site2} = \mu_{site3} = \mu_{site4} =\mu_{site5}$

However, we may have prior expectations that the children growing up in an advantaged site will have a higher score than children growing up in a disadvantaged site. Our informative hypothesis is:

$H_i: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$ -->

<!-- Because it is very improbable that $\beta_{grp}$ is exactly equal to 0, more information can be won by testing $H_i$. -->
<!-- We can test $H_i$ by running a linear regression on the data and passing the `lm` object with the specific hypotheses to `bain`.  -->
<!-- For a single study, such as the study that investigated the Netherlands, the Bayes factor would be obtained as follows: -->

```{r, echo = FALSE, eval = FALSE}
# load bain package
library(bain)

# specify model, allow meaningful zero-point
fit <- lm(soc ~ grp, data = NLa)

#specify Hypotheses to evaluate
hypotheses <- "grp > .1"

fit_bain <- bain(fit, hypothesis = hypotheses)
print(fit_bain)

```
<!-- SESAMESIM DATA: The output of `bain` seems to support neither hypothesis. The `fit` for both hypotheses is approximately zero which means that almost none of the data are in agreement with the restrictions the specified hypotheses pose. The `BF.c`, which is the Bayes Factor of the hypothesis versus its complement, is for both hypotheses close to zero. This implies that the complement of both hypotheses is $\frac{1}{BF.c}$ times more likely than the passed hypotheses. Finally, `PMPc`, the Posterior Model Probabilities of the specified hypotheses with $H_c$ included, is also close to zero for both specified hypothesis. This means that *out of* the specified hypotheses and their complement, the probability of $H_c$ being true is 99.9%. -->

### Aggregating evidence across studies

```{r, eval = FALSE, echo = FALSE, results = resultaction}
# Specify the models for DK and US
model_dk <- "
fam =~ fam_1 + fam_2 + fam_3
con =~ sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
con ~ beta * fam"
model_us <- "
fam =~ fam_1 + fam_2 + fam_3
con =~
secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 + secs_soc_6 + secs_soc_7 +
secs_eco_1 + secs_eco_2 + secs_eco_3 + secs_eco_4 + secs_eco_5
con ~ beta * fam"

# Estimate the model in lavaan
results_dk <- sem(model = model_dk, data = DK)
results_us <- sem(model = model_us, data = US)

# Bind the models into a list
results <- list(results_nl, results_dk, results_us)
# Test the hypothesis that the effect size labeled 'beta' is positive
bf_tot <- pbf(results, hypothesis = "beta > .1", standardize = TRUE)
```

As mentioned before, suitable data were collected to evaluate the substantive hypothesis in three countries.
There are differences between countries that prevent analyzing these data as a multilevel model, however.
For instance, conservatism was measured using different scales.
This is an appropriate situation to use the PBF to aggregate evidence across countries.
Below, we estimate a latent regression model for the remaining two countries, taking care to use the same label for the parameter of interest in all samples.
Then, we bind all three SEM-models in a list, and call PBF to evaluate the hypothesis of interest on all models and aggregate the evidence.
As the BF in all three samples is positive, the resulting PBF is very large.
We can thus conclude that the central hypothesis receives overwhelming support across samples.

```{r, eval = TRUE, echo = TRUE}
# Specify the models for DK and US
model_dk <- "
fam =~ fam_1 + fam_2 + fam_3
con =~
sepa_soc_1 + sepa_soc_2 + sepa_soc_3 + sepa_soc_4 + sepa_soc_5 +
sepa_eco_1 + sepa_eco_2 + sepa_eco_3 + sepa_eco_4 + sepa_eco_5
con ~ beta * fam"
model_us <- "
fam =~ fam_1 + fam_2 + fam_3
con =~
secs_soc_1 + secs_soc_2 + secs_soc_3 + secs_soc_4 + secs_soc_5 +
secs_soc_6 + secs_soc_7 +
secs_eco_1 + secs_eco_2 + secs_eco_3 + secs_eco_4 + secs_eco_5
con ~ beta * fam"

# Estimate the model in lavaan
results_dk <- sem(model = model_dk, data = DK)
results_us <- sem(model = model_us, data = US)

# Bind the models into a list
results <- list(results_nl, results_dk, results_us)
# Test the hypothesis that the effect size labeled 'beta' is positive
pbf(results, hypothesis = "beta > .1", standardize = TRUE)
```

<!-- In the first use-case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`. -->
<!-- `pbf()` supports lists of all objects that `bain()` also supports. All supported types are specified in bain vignette *REF*. One exception is the use of named vectors, which will be covered later. -->
<!-- ```{r} -->
<!-- lm_sample1 <- lm(postnumb~site-1, sample1) -->
<!-- lm_sample2 <- lm(postnumb~site-1, sample2) -->
<!-- lm_sample3 <- lm(postnumb~site-1, sample3) -->

<!-- lm_allsamples <- list(lm_sample1, lm_sample2, lm_sample3) -->

<!-- pbf(lm_allsamples, hypotheses) -->
<!-- ``` -->

<!-- The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually. The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low. When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually. For sample 3 the specified hypotheses are even more likely than their complement. This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation. If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation. This leads to a more extreme combined evaluation than any individual evaluation. Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true. -->

<!-- Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$.  -->
<!-- To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the pbf calculated with `pbf()`.  -->
<!-- In this example, data to evaluate $H_i$ is available for three countries. -->

<!-- In the first use-case of `pbf()`, an `lm()` is ran for each country.  -->
<!-- All `lm` objects are stored in a list which is passed to `pbf()`. -->
<!-- `pbf()` supports lists of all objects that `bain()` also supports.  -->
<!-- All supported types are specified in the `bain` vignette.  -->
<!-- One exception is the use of named vectors, which will be covered later. -->
<!-- ```{r} -->
<!-- fit_US <- lm(soc ~ grp + her, data = USa) -->
<!-- fit_DK <- lm(soc ~ grp + her, data = DKa) -->
<!-- fit_NL <- lm(soc ~ grp + her, data = NLa) -->

<!-- all_fits <- list(fit_US, fit_DK, fit_NL) -->

<!-- pbf(all_fits, hypotheses) -->
<!-- ``` -->

### Using bain objects

The `pbf()` function also accepts multiple `bain` objects.
This makes it possible to, for example, evaluate different sets of hypotheses on different data sets before using the resulting `bain` objects to aggregate the evidence for all common hypotheses across datasets.
The example below illustrates this use case.
As before, all analyses share one hypotheses in common ($H_i: \beta_{fam} > .1$),
but the Dutch sample now contains a sample-specific hypothesis regarding the effect of group morality, namely that $\beta_{grp} < .1$.
The `pbf()` function is called on a list of `bain` objects.
Note that, in this case, `pbf()` does not require an argument `hypothesis`, as the hypotheses are contained in the `bain` objects.

```{r, echo = TRUE, eval = TRUE, results = resultaction}
# Add the additional predictor to the model, label the effect beta2
model_nl <- c(model_nl, "group =~ grp_1 + grp_2 + grp_3
                         con ~ beta2 * group")

# Estimate the model in lavaan
results_nl <- sem(model = model_nl, data = NL)

# Obtain BF for each sample; note that the Dutch sample has two hypotheses
bf_nl <- bain(results_nl, hypothesis = "beta > .1;
                                        beta2 < .1", 
              standardize = TRUE)
bf_dk <- bain(results_dk, "beta > .1")
bf_us <- bain(results_us, "beta > .1")

# Bind bain objects into a list
bfs <- list(bf_nl, bf_dk, bf_us)

# Call pbf on that list
pbf(bfs)
```

As can be seen, the results are equivalent to the results in the previous example.
The sample-specific hypothesis has been left out, and common hypotheses are retained and aggregated.
If there are no common hypotheses across all objects, `pbf()` throws an error.

```{r, echo = FALSE, eval = FALSE}
bains_different_hyps <- list(
  bain(fit_US, "grp = 0"),
  bain(fit_DK, "grp > .1"),
  bain(fit_NL, "grp < .1")
)

tryCatch({
  pbf(bains_different_hyps)},
  error = function(e){message(e)}
      )

```
<!--The function returns an error stating that the samples have no hypothesis in common and thus the PBF cannot be calculated.-->

### Using sufficient statistics

A third use-case occurs when the raw data from different samples are not available.
This may happen, for example, when aggregating findings from the published literature (similar to meta-analysis).
In this case, one can use the default interface of `bain`, as explained in [@hoijtink2019tutorial].
This function requires four arguments: A named vector of parameter estimates, their asymptotic covariance matrix, the original sample size, and the number of within-group and between-group parameters.
Note that, when analyzing a single parameter per sample, the standard error is sufficient to construct the asymptotic covariance matrix.
Thus, this method can be applied to data that have been prepared for classic meta-analysis (effect sizes and their sampling variances).
Importantly, unlike meta-analysis, the present method is suitable for conceptual replications.
It does not require uniform effect size measures across studies.
The example below illustrates how to aggregate evidence for one hypotheses across three studies that each used different methods.
<!-- CJ: What I'd like to do here is show how to aggregate a t-test, a regression coefficient and a correlation coefficient -->

The present use case evaluates the following hypothesis:
*There is a positive association between family morality and political conservatism*.
This conceptual hypothesis is evaluated differently in the three samples, resulting in three different types of statistics and distinct sample-specific hypotheses:

1. A t-test was performed using the NL data; using Cohen's D gives $H_i^{NL}: \delta_{conservative > liberal} > 0$, where $\delta$ is the mean difference between groups.
1. A bivariate regression coefficient was calculated using the DK data, giving $H_i^{DK}: \beta_{fam} > 0$
1. A correlation coefficient was calculated using the US data, giving $H_i^{US}: \rho_{fam,con} > 0$, where $\rho$ is the correlation between family morality and conservatism.

Note that we intentionally manipulate the data to illustrate these different analyses;
for example, we compute mean scale scores and dichotomize the continuous conservatism scale to conduct a t-test.
We do not advocate these practices for applied research.

First we obtain the relevant parameter estimates and their sampling variances,
which allows us to evaluate the specific hypotheses in `bain`:

```{r eval = TRUE, echo = TRUE}
# Create mean scale scores
NL <- data.frame(
  family = rowMeans(NL[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(NL[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
                               "sepa_eco_5")]))
DK <- data.frame(
  family = rowMeans(DK[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(DK[c("sepa_soc_1", "sepa_soc_2", "sepa_soc_3",
                               "sepa_soc_4", "sepa_soc_5", "sepa_eco_1",
                               "sepa_eco_2", "sepa_eco_3", "sepa_eco_4",
                               "sepa_eco_5")]))

US <- data.frame(
  family = rowMeans(US[c("fam_1", "fam_2", "fam_3")]),
  conservative = rowMeans(US[c("secs_soc_1", "secs_soc_2", "secs_soc_3",
                               "secs_soc_4", "secs_soc_5", "secs_soc_6",
                               "secs_soc_7", "secs_eco_1", "secs_eco_2",
                               "secs_eco_3", "secs_eco_4", "secs_eco_5")]))

# NL: Conduct t-test using Cohen's D
NL$group <- cut(NL$conservative, breaks = 2,
                labels = c("liberal", "conservative"))
sample_sizes <- table(NL$group)
sds <- tapply(NL$family, NL$group, sd)
pooled_sd <- sqrt(sum((sample_sizes - 1) * sds) / (sum(sample_sizes) - 2))
NL_est <- diff(tapply(NL$family, NL$group, mean)) / pooled_sd
NL_var <- (sum(sample_sizes) / prod(sample_sizes)) +
  (NL_est^2 / (2*sum(sample_sizes)))

# DK: Conduct bivariate regression
DK_fit <- lm(conservative ~ family, data = DK)
DK_est <- coef(DK_fit)["family"]
DK_var <- vcov(DK_fit)["family", "family"]

# US: Correlation coefficient
US_est <- cor(US)[1, 2]
US_var <- (1 - US_est^2)^2 / (nrow(US) - 1)

# Name the estimates so hypotheses will be the same
names(NL_est) <- names(DK_est) <- names(US_est) <- "parameter"
```

Then, we use `bain.default()` to evaluate the central hypothesis on each parameter estimate. 
The `pbf()` function can be called on a list of the resulting bain objects.

```{r eval = TRUE, echo = TRUE, results = resultaction}
# Use bain.default() to obtain BF for the central hypothesis 
NL_bain <- bain(x = NL_est, 
                Sigma = matrix(NL_var, 1, 1),
                n = nrow(NL),
                hypothesis = "parameter > 0",
                joint_parameters = 1)
DK_bain <- bain(x = DK_est,
                Sigma = matrix(DK_var, 1, 1),
                n = nrow(DK),
                hypothesis = "parameter > 0",
                joint_parameters = 1)
US_bain <- bain(x = US_est,
                Sigma = matrix(US_var, 1, 1),
                n = nrow(US),
                hypothesis = "parameter > 0",
                joint_parameters = 1)

# Aggregate evidence using pbf()
pbf(list(US_bain, DK_bain, NL_bain))
```

The results suggest substantial evidence for the hypothesis that there is a positive association between family morality and political conservatism.
Although each study used a different method to assess this hypothesis,

```{r, eval = FALSE, echo = FALSE}
get_estimates <- function(sample){
  number_sites <- length(levels(sesamesim$site))

  sample_size <- nrow(sample)

  site_estimates <- sapply(1:number_sites, function(i){
    rbind(
      mean(sample[sample['site'] == i, "postnumb"]),
      sd(sample[sample['site'] == i, "postnumb"])
      )
    })

  est_names <- sprintf("site%d", 1:number_sites)
  site_means <- site_estimates[1,]
  names(site_means) <- est_names

  covmat <- diag((site_estimates[2,] / sqrt(sample_size))^2)
  dimnames(covmat) <- list(est_names, est_names)
  return(list(means = site_means, covmat = covmat, sample_size = sample_size))
}

ests_sample1 <- get_estimates(sample1)
ests_sample2 <- get_estimates(sample2)
ests_sample3 <- get_estimates(sample3)

```

```{r eval = FALSE, echo = FALSE, results = resultaction}
bain_sample1_ests <- bain(x = ests_sample1[['means']], 
                          Sigma = ests_sample1[['covmat']], 
                          n = ests_sample1[['sample_size']], 
                          hypothesis = hypotheses)

bain_sample2_ests <- bain(x = ests_sample2[['means']], 
                          Sigma = ests_sample2[['covmat']], 
                          n = ests_sample2[['sample_size']], 
                          hypothesis = hypotheses)

bain_sample3_ests <- bain(x = ests_sample3[['means']], 
                          Sigma = ests_sample3[['covmat']], 
                          n = ests_sample3[['sample_size']], 
                          hypothesis = hypotheses)

bain_allsamples_ests = list(bain_sample1_ests, bain_sample2_ests, bain_sample3_ests)
pbf(bain_allsamples_ests)
```

<!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use-case, or is something off in the script? -->
<!-- CJ: We should ask Herbert about this, but this example won't make it into this paper. If you want, you can send Herbert a reproducible example of the 'problem'!-->

# Conclusion

<!-- Researchers need to choose a method that best suits their research question, -->
 <!-- the study highlights the importance of carefully considering the research question when selecting a method for meta-analysis, and the potential trade-offs that may exist between sensitivity and specificity when using different methods. -->

In conclusion, this study evaluated the performance of the product Bayes factor as a method for evidence synthesis,
and compared it against other commonly used evidence synthesis methods under different simulation conditions.
Compared to the other methods,
PBF had the highest overall accuracy.
This was primarily due to its greater sensitivity.
However, PBF had lower specificity than all other algorithms, suggesting a trade-off between sensitivity and specificity.
The other algorithms showed ceiling effects in specificity, limiting their sensitivity.
The performance of the PBF was most strongly affected by sample size,
We introduced a user-friendly implementation of the PBF in the `bain` R-package,
and demonstrated its use with various analysis techniques in R,
as well as with sufficient statistics that are already routinely coded for meta-analysis (i.e., effect sizes and their sampling variance).
This means that researchers can now use the PBF to aggregate evidence in situations where classic meta-analytic methods are less suitable.
For example, when one informative hypothesis has been evaluated in several replication studies,
but these replication studies are quite heterogeneous because they sample from different populations and use different methods or analysis techniques.
Especially when the number of replication studies is too small to adequately account for these sources of between-study heterogeneity,
the PBF may be a useful method to aggregate evidence for the common informative hypothesis.
Researchers should be aware that the PBF trades off increased sensitivity for decreased specificity,
and that it addresses a different research question than other research synthesis methods.
This highlights the importance of careful interpretation of the results,
and consideration of the research question when selecting an aggregation method.
In sum,
our results suggest that PBF is a useful evidence synthesis method,
which is now broadly accessible due to its inclusion in the `bain` R-package.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup